Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2017-06-21 code/rmarkdown/article.Rmd
inserted line numbers

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Earth and Environmental Sciences, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  JOH performed the analyses. All authors developed the study and drafted the manuscript.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue. However, genetic variation, the raw material for evolution, is difficult to capture in conservation planning, not least because genetic data require considerable resources to obtain and analyze. Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation. We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to prioritize protected area acquisitions. First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation. Second, we showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Finally, we discovered that solutions based only on surrogate information secured substantial amounts of adaptive and neutral genetic variation. Our work paves the way for widespread integration of surrogates for genetic variation into conservation planning.
significance: |
  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different genetic types. However, genetic data are expensive and time consuming to obtain. Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning. This means that conservation planners can, with some confidence, design protected area systems to represent intra-specific genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

Although the strength of natural selection is a continuous force, genetic variation is often classified as adaptive or neutral (reviewed in [@r480]). Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg. [@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg. [@r500]). By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]). In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]). Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).

Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])--without needing to utilize genetic data to achieve this. Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg. [@r202; @r2]). Specifically, by conserving individuals in a broad range of environmental conditions, we might expect to capture individuals with a diverse set of local adaptations, and overall capture a large proportion of the adaptive genetic variation present in the species. However, the effectiveness of this approach remains unverified. Neutral genetic variation arises from a reduction in gene flow between populations. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]). Potential surrogates for neutral genetic variation have been based on variables that predict the level of connectivity between different areas. For instance, according the isolation by distance model (IBD; [@r259]), populations located further apart are predicted experience less gene flow, and in turn, share fewer neutral loci. Based on this idea, spreading conservation priorities evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]). However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg. land use; [@r497])--the typical situation in most conservation planning exercises.

@@

<!-- summary of methods -->
We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species. We used two outlier detection methods to identify loci under selection. These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1). Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive ($K = `r unique(na.omit(adaptive.k))`$ for all species; Table S2) and neutral genetic variation ($`r min(neutral.k)` \leq K \leq `r max(neutral.k)`$ for all species; Table S1). The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (eg. _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.

@@

<!-- surrogacy analysis -->
For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritizations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation that each prioritization captured. We then repeated this process, and calculated the proportion of variation in geographic position and neutral genetic variation they sampled. The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figures 1, S28 and S29). The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). _Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4). After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogate variables.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the median, and the 25th and 75th percentiles. Whiskers show the data at 1.5 $\times$ the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(Prioritisation, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame()

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2). Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors ([@r505; @r497]). For example, some species can maintain relatively high connectivity between distant populations (eg. wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (eg. anthropogenically modified land [@r497]). Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.

@@

<!-- adaptive genetic variation secured in prioritizations-->
Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2). For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg. _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively). As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample. This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations. Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a). Overall--and contrary to our original hypothesis--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. Firstly, the size of the planning units we used (approx. 23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorporate data on topography may better describe connectivity in the study area. However, such distances often require species-specific scaling (eg. [@r497]), and so cannot easily be utilized in multi-species planning exercises. Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to describe genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 3a). All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}). This data set has been used to explore patterns of adaptive (eg. [@r455]) and neutral genetic variation (eg. [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]). Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 $\times$ 25 km; Figure S31). Project members visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found. They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid. We used this data because it comprised comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.

@@

To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35). Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]). The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]). The advantage of these methods--in contrast with environmental association analyses--is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it. Loci identified by both outlier detection methods were treated as adaptive and the remainder were treated as neutral. To minimize false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

@@

The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r487]). To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis. For each of the 27 species, we first imputed missing data by replacing missing values with the average frequency among conspecifics, then ran a PCA over the loci matrix, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci ( `r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci). We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).

@@

We used the unreliable representative and adequate prioritization (URAP) formulation to generate solutions ([@r465]). This formulation can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values--because they secured very little genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package; [@r465]).

>>>>

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Earth and Environmental Sciences, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  JOH performed the analyses. All authors developed the study and drafted the manuscript.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue. However, genetic variation, the raw material for evolution, is difficult to capture in conservation planning, not least because genetic data require considerable resources to obtain and analyze. Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation. We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to prioritize protected area acquisitions. First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation. Second, we showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Finally, we discovered that solutions based only on surrogate information secured substantial amounts of adaptive and neutral genetic variation. Our work paves the way for widespread integration of surrogates for genetic variation into conservation planning.
significance: |
  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different genetic types. However, genetic data are expensive and time consuming to obtain. Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning. This means that conservation planners can, with some confidence, design protected area systems to represent intra-specific genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship.
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

Although the strength of natural selection is a continuous force, genetic variation is often classified as adaptive or neutral (reviewed in [@r480]). Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg. [@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg. [@r500]). By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]). In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]). Thus optimally sited protected area networks would aim to represent patterns of both adaptive and neutral genetic variation ([@r67; @r3; @r43]).

Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])---without needing to utilize genetic data to achieve this. Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg. [@r202; @r2]). Specifically, by conserving individuals in a broad range of environmental conditions, we might expect to capture individuals with a diverse set of local adaptations, and overall capture a large proportion of the adaptive genetic variation present in the species. However, the effectiveness of this approach remains unverified. Neutral genetic variation arises from a reduction in gene flow between populations. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]). Potential surrogates for neutral genetic variation have been based on variables that predict the level of connectivity between different areas. For instance, according the isolation by distance model (IBD; [@r259]), populations located further apart are predicted experience less gene flow, and in turn, share fewer neutral loci. Based on this idea, spreading conservation priorities evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]). However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg. land use; [@r497])---the typical situation in most conservation planning exercises.

@@

<!-- summary of methods -->
We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species. We used two outlier detection methods---that solely used genetic data and did not use environmental data---to identify loci under selection. These methods returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD; Table S1). Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive ($K = `r unique(na.omit(adaptive.k))`$ for all species; Table S2) and neutral genetic variation ($`r min(neutral.k)` \leq K \leq `r max(neutral.k)`$ for all species; Table S1). The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S1--S27) generally showed substantial spatial auto-correlation (eg. _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figure S36), suggesting that the environmental and geographic variables have potential to be effective surrogates for genetic variation.

@@

<!-- surrogacy analysis -->
For each species, we generated a suite of `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates, big.mark=',', scientific=FALSE)` random prioritizations, and calculated the proportion of environmental variation and adaptive (if present) genetic variation that each prioritization captured. We then repeated this process, and calculated the proportion of variation in geographic position and neutral genetic variation they sampled. The environmental and geographic variables were moderately effective predictors for the genetic variation represented by the randomly generated prioritizations (Figures 1, S28 and S29). The relationship between the proportion of genetic and surrogate variation secured in the prioritizations varied among the species (species $\times$ proportion of environmental variation interaction term: ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$; species $\times$ proportion of geographic variation interaction term: ${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P =$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). _Post-hoc_ analyses showed that these relationships were positive for all species (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S4). After establishing that the the proportion of surrogate variation secured in a prioritization also predicted the proportion of genetic variation it secured, our final step was to determine if prioritizations were more effective when generated using targets based on the surrogate variables.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the median, and the 25th and 75th percentiles. Whiskers show the data at 1.5 times the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(Prioritisation, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame()

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic variables are effective surrogates for most species considered here (Figure 2). Despite this, geographic distance was a suprisingly poor surrogate for neutral genetic variation for a few species (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where spatial genetic structure is complicated by additional factors ([@r505; @r497]). For example, some species can maintain relatively high connectivity between distant populations (eg. wind dispersing plants; [@r451]), and in some places can be disrupted by landscape features (eg. anthropogenically modified land [@r497]). Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.

@@

<!-- adaptive genetic variation secured in prioritizations-->
Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2). For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg. _Carex sempervirens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively). As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample. This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations. Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation---they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a). Overall---and contrary to our original hypothesis---it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. Firstly, the size of the planning units we used (approx. 23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorporate data on topography may better describe connectivity in the study area. However, such distances often require species-specific scaling (eg. [@r497]), and so cannot easily be utilized in multi-species planning exercises. Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to describe genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time. Furthermore, even with modern population genomic approaches, a survey can at best hope to identify markers that are linked to functionally adaptive variants.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 3a). All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic-surrogates}). This data set has been used to explore patterns of adaptive (eg. [@r455]) and neutral genetic variation (eg. [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]). Data were collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 $\times$ 25 km; Figure S31). Project surveyors visited every second grid cell, and for each species, they collected samples from three individuals if any individuals were found. They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid. We used these data because they comprised comparable genetic information for a range of species with different evolutionary and life histories collected using a standardized sampling scheme.

@@

To investigate the effectiveness of our putative surrogates, we first needed to identify which of the sampled loci were adaptive (Figure S35). Following best practice, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]). The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]). The advantage of these methods---in contrast with environmental association analyses---is that they do not use environmental data, which would have introduced an element of circularity into our analysis and invalidated it. Loci identified by both outlier detection methods were treated as adaptive and the remainder were treated as neutral. To minimize false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

@@

The second outlier detection method involved fitting PCAs to identify outlier loci (implemented in the \texttt{pcadapt R} package; [@r487]). To enable comparisons between the two outlier detection methods, we used the same individuals in this analysis as in the \texttt{BayeScan} analysis. For each of the 27 species, we first imputed missing data by replacing missing values with the average frequency among conspecifics, then ran a PCA over the loci matrix, and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*100` % of the variation in loci). We then computed _q_-values using Mahalanobis distances, and used the same false discovery rate used in the \texttt{BayeScan} analysis to identify loci under selection (using the \texttt{qvalue R} package; [@r491]).

@@

We used the \texttt{raptr R} package to generate solutions ([@r465]). This toolkit can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values---because they secured very little genetic variation---were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality.
